---
title: "Download an ILSA dataset"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Download}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup,echo=FALSE}
library(ILSAmerge)
```

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE
)
```



## Available datasets

Currently [`ILSAmerge`](https://cran.r-project.org/package=ILSAmerge) is able
to download several ILSA datasets, including: 'PISA', 'TALIS', 'TIMSS', 
'TIMSS Advanced', 'PIRLS', 'ICCS', 'ICILS', 'CIVED', 'REDS', 'RLII', and 'SITES.'

The links of these datasets are allocated in 
[GitHub](https://raw.githubusercontent.com/dopatendo/ILSAmerge/refs/heads/main/data/ILSAlinks.csv),
so links can be corrected if they are down.

With a small script, we can check which datasets are available to download:
```{r, eval = TRUE}
ilsaliks <- "https://raw.githubusercontent.com/dopatendo/ILSAmerge/refs/heads/main/data/ILSAlinks.csv"
ilsas <- unique(read.csv(ilsaliks)[,1:2])
ilsas <- ilsas[!ilsas$Name%in%"Other",]
ilsasU <- sort(unique(ilsas$Name))

for(i in 1:length(ilsasU)){
  cat(paste0(ilsasU[i],": ",
             paste0(ilsas$Year[ilsas$Name%in%ilsasU[i]],collapse = ", "),".\n"))
}
```


## Downloading the datasets for a particular study

Depending on the study datasets are uploaded in 1 or more .zip files. That is why
`ILSAdownload()` will handle by itself how many and which .zip files to download.

As an example, we let us say we want to download the data of 'RLII' 1991. Then,
at least we need to opt for three arguments; `study`, `year`, and `outputdir`:
```{r, eval = FALSE}
ILSAdownload(study = "RLII", year = 1991, outputdir = tempdir())
```
```{r, eval = TRUE, echo=FALSE}
try(ILSAdownload(study = "RLII", year = 1991, outputdir = tempdir()),silent = TRUE)
```

As we can see, some studies will require that you agree with their license 
agreements. If that is the case, you will be required to answer if you agree 
or not. In order to continue, you will need to agree. If you know that you
will agree with the study terms, then you can also avoid this prompt via
`agreeLicense`:
```{r, eval = TRUE}
ILSAdownload(study = "RLII", year = 1991, outputdir = tempdir(),
             agreeLicense = TRUE)
```

## Unzipping

This function can also unzip downloaded files using `unzip = TRUE`:

```{r, eval = FALSE}
ILSAdownload(study = "RLII", year = 1991, outputdir = tempdir(),
             unzip = TRUE)
```

## Avoid printing

It is possible to avoid any printing by `quiet = TRUE`, but you will still
be prompted to agree. To avoid both prints you can use both arguments:

```{r, eval = FALSE}
ILSAdownload(study = "RLII", year = 1991, outputdir = tempdir(),
             quiet = TRUE, agreeLicense = TRUE)
```

## Downloading time

Timeout times for downloading are controlled by `maxtime`, which you can set
for a desired number of seconds:

```{r, eval=FALSE}
ILSAdownload(study = "RLII", year = 1991, outputdir = tempdir(),
             maxtime = 60)
```
