---
title: "Identify ILSA files"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Identify ILSA files}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE
)
```

```{r setup, echo = FALSE}
library(ILSAmerge)
```


It is common that ILSA files come an unaggregated format, i.e., separated by
country. If that is the case [`ILSAmerge`](https://cran.r-project.org/package=ILSAmerge)
is able to identify which files should be merged together.

If we want know how many files were downloaded and into which many files
they can be merged, we can use `ILSAfile.info()` targeting the folder with
the unzipped data:

```{r}

ILSAdownload(study = "REDS", year = 2021, outputdir = tempdir(),
             unzip = TRUE, quiet = TRUE, agreeLicense = TRUE)

ILSAfile.info(inputdir = file.path(tempdir(),"REDS2021_IDB_SPSS/Data"))

```

As we can see, for 'REDS' 2021 we have 3 file 'types' or populations: `BCGV1` 
(schools), `BSGV1` (students), and `BTGV1` (teachers). This will become important
when we need to work with only one population.
